---
title: "CTF446_scRNA_ver2"
format: html
editor: source
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
set.seed(8578)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  results = "hide"
)


library(dplyr)
library(Seurat)
library(patchwork)
library(knitr)
library(ggplot2)
library(cowplot)
library(tidyverse)
library(data.table)
library(igraph)
library(ggraph)
library(clustree)
library(HGNChelper)
library(presto)
library("SeuratExtend")
```

#1. QC, filtering, and normalisation
##1.1 Data import

```{r load data}

## Load data (Rhapsody already a Seurat object with metadata from ST and TCR)


gt <- 
  readRDS(file = "CTF446_updatedfiles/CTF446-D11-GT-run4_Seurat.rds")

##make a list of cell label for velocyto and export

# cell_list <- Cells(gt)
# 
# writeLines(cell_list, con = "CTF446_updatedfiles/output/cell_list.txt", sep = "\n", useBytes = FALSE)

# update Seurat object

gt[["RNA"]] <- as(object =  gt[["RNA"]], Class = "Assay5")
#
#get bioproduct names for RNA
rna_name <- rownames(gt)
head(rna_name)

#get RNA meta data

rna <- GetAssayData(gt[["RNA"]], features = rna_name )

#recreate seurat object

gt[["RNA"]] <- CreateAssay5Object(counts = rna)

head(gt[["RNA"]]$counts)

#remove unidentified genes

gt <- gt[!rownames(gt) %>% stringr::str_detect("Gm\\d{1,}|\\d{1,}Rik|ENSMUSG"), ]

```

## Remove multiplet and undetermined

```{r multiplet and undetermined}

gt@meta.data %>%
  group_by(Sample_Name) %>%
  tally(name = "cell_count") %>%
  ggplot(aes(x = Sample_Name, y = cell_count, fill = Sample_Name)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(y = cell_count, 
                label = cell_count,
                vjust = -0.5, size = 2.5), show.legend = F) +
  scale_x_discrete(guide = guide_axis(angle = 45)) + 
  theme_classic()

#remove 

gt <- subset(gt, subset = Sample_Name %in% c("Multiplet","Undetermined"), invert = T)



```

##1.2 QC


```{r explore RNA content}

# visualize QC metrics
gt[["percent.mt"]] <- PercentageFeatureSet(gt, pattern = "^mt-")
head(gt@meta.data, 5)

VlnPlot(gt, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

p1 <- VlnPlot(gt, features = c("nFeature_RNA"))+ geom_hline(yintercept = 3000 , linetype = "dashed", color = "red") + theme(legend.position = "none")
          
p2 <- VlnPlot(gt, features = "percent.mt", y.max = 25) + geom_hline(yintercept = 12.5 , linetype = "dashed", color = "red") + theme(legend.position = "none")

plot_grid(p1, p2, ncol = 2)

plot1 <- FeatureScatter(gt, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(gt, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2

#visualisation by samples

gt@meta.data %>% 
  ggplot(aes(color=Sample_Name, x=nCount_RNA, fill= Sample_Name)) + 
  geom_density(alpha = 0.2) + 
  scale_x_log10() + 
  theme_classic() +
  ylab("Cell density") +
  ggtitle("nCount_RNA")

gt@meta.data %>% 
  ggplot(aes(color=Sample_Name, x=nFeature_RNA, fill= Sample_Name)) + 
  geom_density(alpha = 0.2) + 
  scale_x_log10() + 
  theme_classic() +
  ylab("Cell density") +
  ggtitle("nFeature_RNA")

gt@meta.data %>% 
  ggplot(aes(color=Sample_Name, x=percent.mt, fill=Sample_Name)) + 
  geom_density(alpha = 0.2) + 
  theme_classic() +
  ggtitle("percent.mt")

gt@meta.data %>% 
  ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  geom_point() + 
  scale_colour_gradient(low = "gray90", high = "black") +
  stat_smooth(method=lm) +
  scale_x_log10() + 
  scale_y_log10() + 
  theme_classic() +
  geom_vline(xintercept = 700, colour = "red") +
  geom_hline(yintercept = 400, colour = "red") +
  facet_wrap(~Sample_Name) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

"The data is fitted with a linear regression line. We expect when UMI counts increase, gene counts also increase.

Upper right quadrant: Cells with high gene counts and UMI counts. We generally consider these cells to be good-quality cells.

Bottom right quadrant: Cells with low gene counts and UMI counts. We generally consider these cells are low-quality cells.

Darker cells: Cells with high mitochondrial percentage. Many of these cells have low gene counts at the same time. This may indicate damaged/dying cells whose cytoplasmic mRNA leaks out through a broken membrane, so only mRNA located in the mitochondria is still conserved.
"

```


```{r filtering}

"From the QC plot I will set a mild threshold (mt = 12,5 %)"

gt_filtr <- subset(gt, subset = nFeature_RNA > 400 & nFeature_RNA < 3000 & percent.mt < 12.5 & nCount_RNA > 700 )

VlnPlot(gt_filtr,features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, alpha = .2 )

#assess filtration
"Before filtering"
gt@meta.data %>%
  group_by(Sample_Name) %>%
  dplyr::summarise(cell_number = length(Sample_Name))

"After filtering"
gt_filtr@meta.data %>%
  group_by(Sample_Name) %>%
  dplyr::summarise(cell_number = length(Sample_Name))
```

#Scaling, norm and findvariable features

```{r}
options(future.globals.maxSize= 600*1024^2)
# run sctransform
gt_norm <- SCTransform(gt_filtr)

#run normalisation and scaling for RNA slot to use in downstream analysis such as celltypist, singleR, DE
#SCT data use for clustering, UMAP, and integration

gt_norm <-
  NormalizeData(gt_norm, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")


all.genes <- rownames(gt_norm)
gt_norm <- ScaleData(gt_norm, features = all.genes, assay = "RNA")


```


##1.6 Linear dimensial reduction
```{r PCA}
set.seed(8578)
gt_norm <- RunPCA(gt_norm, features = VariableFeatures(object = gt_norm))

ElbowPlot(gt_norm, ndims = 50)



#visualise

print(gt_norm[["pca"]], dims =1:5, nfeatures = 5)

VizDimLoadings(gt_norm, dims = 1:2, reduction = "pca")

DimPlot(gt_norm, reduction = "pca") + NoLegend()

DimHeatmap(gt_norm, dims= 1, cells = 500, balanced = TRUE)

DimHeatmap(gt_norm, dims = 1:15, cells = 500, balanced = TRUE)


# Determine percent of variation associated with each PC
pct <- gt_norm[["pca"]]@stdev / sum(gt_norm[["pca"]]@stdev)*100

# Calculate cumulative percents for each PC
cumu = cumsum(pct)


# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5 and variation between consecutive PCs is less than 0.1 %
co1 <- which(cumu > 90 & pct < 5)[1]

co1

# Determine the difference between variation of PC and subsequent PC
co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

# last point where change of % of variation is more than 0.1%

co2

#minimum of the two calculation
pcs <- min(co1, co2)
pcs

"Number of PCs chosen for downstream analysis: 13 (based on the elbow plot and calculation)"

# Create a dataframe with values
plot_df <- data.frame(pct = pct, 
           cumu = cumu, 
           rank = 1:length(pct))

# Elbow plot to visualize 
  ggplot(plot_df, aes(cumu, pct, label = rank, color = rank > pcs)) + 
  geom_text() + 
  geom_vline(xintercept = 90, color = "grey") + 
  geom_hline(yintercept = min(pct[pct > 5]), color = "grey") +
  theme_bw()


```
2. Downstream analysis
##2.1 Clustering
```{r Cluster}
set.seed(8578)

#cluster cells using KNN algorithm

gt_norm <- FindNeighbors(gt_norm,  reduction = "pca", dims = 1:13)

#UMAP use only the top 13 PCs

gt_norm <- RunUMAP(gt_norm, reduction = "pca", dims = 1:13)


DimPlot(gt_norm, reduction = "umap", split.by = "Sample_Name")


DimPlot(gt_norm, reduction = "tsne", split.by = "Sample_Name")
```


```{r upload old gt norm}
library("svglite")


gt_norm <- FindClusters(gt_norm, resolution = c(0.05, 0.1, 0.2, 0.3),
                         verbose = F)

# gt_norm <- FindClusters(gt_norm, resolution = 0.1,verbose = F)

#look at metadata for each resolution how many clusters
# gt_norm@meta.data %>%
#   dplyr::select(contains("RNA_snn")) %>%
#   map_int(~ unique(.x) %>% length)

"At resolution 0.1 we see nodes with multiple incoming edges - this is a good indication of overclustering.
Here it might be good to chose the 0.05 as the resolution"

#visualise clusters at different resolution to determine it

clustree(gt_norm, prefix = "SCT_snn_res.")
# 
Idents(object = gt_norm) <- "SCT_snn_res.0.05"

# Idents(gt_norm)
# 
# UMAP visualization
y <- DimPlot(gt_norm, reduction = "umap", label = TRUE, label.size = 6)

y

```

```{r CD4 clusters}

"Remove clusters that are not CD4"
T_cell_marker <- VlnPlot(gt_norm, features = c("Ptprc","Cd4", "Cd3e", "EYFP",  "Cd3g", "Thy1", "Epcam", "Cd8a", "Klrb1c", "Ncr1", "Pecam1"))

T_cell_marker

# ggsave("CTF446_updatedfiles/output/cluster_confirmCD4.svg", plot = T_cell_marker, width = 30, height = 30, units = "cm")

library(scCustomize)

perc.CD4.all <- Percent_Expressing(gt_norm, features = c( "Ptprc","Cd4", "Cd3e", "EYFP",  "Cd3g", "Thy1", "Epcam", "Cd8a", "Klrb1c", "Ncr1", "Pecam1" ), threshold = 1)

perc.CD4.all
# 
# write.csv(perc.CD4.all, "CTF446_updatedfiles/output/perc_Tmarker2.csv", row.names = T)

"Remove cluster 3"
```

# 3 VDJ analysis

```{r}
library(gsl)
library(scRepertoire)
library(data.table)

tcr <- fread("CTF446_updatedfiles/CTF446-D11-GT-run4_VDJ_Dominant_Contigs_AIRR.tsv", sep="\t")  %>%
  dplyr::rename(chain = locus,
         barcode = cell_id,
         reads = umi_count,
         v_gene = v_call,
         d_gene = d_call,
         j_gene = j_call,
         c_gene = c_call,
         cdr3_nt = cdr3,
         cdr3 = cdr3_aa)

st <- fread("CTF446_updatedfiles/CTF446-D11-GT-run4_Sample_Tag_Calls.csv")

#split data by sampletag and T cells by chain
sample1 <- filter(tcr, barcode%in% (filter(st, Sample_Tag == "SampleTag05_mm") %>% .$Cell_Index), chain %in% c("TRB", "TRA", "TRG", "TRD"))
sample2 <- filter(tcr, barcode%in% (filter(st, Sample_Tag == "SampleTag06_mm") %>% .$Cell_Index), chain %in% c("TRB", "TRA", "TRG", "TRD"))
sample3 <- filter(tcr, barcode%in% (filter(st, Sample_Tag == "SampleTag07_mm") %>% .$Cell_Index), chain %in% c("TRB", "TRA", "TRG", "TRD"))
sample4 <- filter(tcr, barcode%in% (filter(st, Sample_Tag == "SampleTag08_mm") %>% .$Cell_Index), chain %in% c("TRB", "TRA", "TRG", "TRD"))

bd <- list(sample1, sample2, sample3, sample4)


combined.tcr <- combineTCR(bd,
                           samples = c("SampleTag05_mm", "SampleTag06_mm", "SampleTag07_mm", "SampleTag08_mm"),
                           removeNA = FALSE,
                           removeMulti = FALSE,
                           filterMulti = FALSE)

```


```{r combine with seurat object}
library(SingleCellExperiment)
```


```{r combine with seurat object}
cell.barcodes <- rownames(gt_norm[[]])

cell.barcodes <- paste0(gt_norm$Sample_Name ,"_", cell.barcodes)
testing <- RenameCells(gt_norm, new.names = cell.barcodes)

comb <- combineExpression(combined.tcr, testing, cloneCall = "gene",  group.by = "sample" ,proportion = TRUE)

DimPlot(comb, group.by = "cloneSize")
```

```{r exclude clusters}

VlnPlot(comb, features = c("Cd4", "Cd3e", "EYFP"))

#filter irrelevant clusters
# Exclude specific clusters
exclude <- c(2)
# Get the clusters that are not in the exclude list
include_clusters <- setdiff(unique(comb@meta.data[["SCT_snn_res.0.05"]]), exclude)

# Subset the Seurat object to include only the desired clusters
comb <- subset(comb, idents = include_clusters)

# # Ensure UMAP has been computed (run this if not already done)
# if (!"umap" %in% Reductions(comb)) {
#   gt_norm1 <- RunPCA(gt_norm, features = VariableFeatures(object = gt_norm))
#   gt_norm1 <- RunUMAP(gt_norm, dims = 1:10)
#}

# UMAP visualization
DimPlot(comb,
        reduction = "umap",
        label = TRUE,
        label.size = 6)



```


# 4. reclustering
```{r recluster}
set.seed(8578)



# Step 2: sctransform the data in the subset (if not already normalized)
CD4_cluster <- SCTransform(comb)

# Step 5: Perform PCA on the subset
CD4_cluster <- RunPCA(CD4_cluster, features = VariableFeatures(object = CD4_cluster))

# Determine percent of variation associated with each PC
pct <- CD4_cluster[["pca"]]@stdev / sum(CD4_cluster[["pca"]]@stdev)*100

# Calculate cumulative percents for each PC
cumu = cumsum(pct)

ElbowPlot(CD4_cluster, ndims = 50)

```


```{r recluster}
# Step 6: cluster cells using KNN algorithm

CD4_cluster <- FindNeighbors(CD4_cluster,  reduction = "pca", dims = 1:10)

#UMAP use only the top 10 PCs

CD4_cluster <- RunUMAP(CD4_cluster, reduction = "pca", dims = 1:10)

DimPlot(CD4_cluster, reduction = "umap", label = TRUE, label.size = 6)


# Step 7: Find neighbors and clusters (reclustering)
CD4_cluster <- FindClusters(CD4_cluster, resolution = c(0.1,0.12,0.14,0.16,0.18,0.2), verbose = F)
clustree(CD4_cluster, prefix = "SCT_snn_res.")

Idents(object = CD4_cluster) <- "SCT_snn_res.0.16"

y <- DimPlot(CD4_cluster, reduction = "umap", label = TRUE, label.size = 6)

y
```

```{r}
###make a piechart of the percentage of each cluster

library(ggplot2)
library(Seurat)
library(dplyr)

# Count cells per cluster
cluster_counts <- CD4_cluster %>% 
  Idents() %>%
  table() %>%
  as.data.frame()

colnames(cluster_counts) <- c("cluster", "count")

# Calculate percentage
cluster_counts <- cluster_counts %>%
  mutate(percent = round(100 * count / sum(count), 1))

# Plot
ggplot(cluster_counts, aes(x = "", y = percent, fill = cluster)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(percent, "%")), 
            position = position_stack(vjust = 0.5),
            color = "white",  # <- This makes the text white
             size = 6
            ) +
  theme_void() +
  theme(legend.title = element_blank())


```

##2.3 Marker identification
```{r markers}

# find markers for every cluster compared to all remaining cells
#by default wilcoxon Rank test
gt.markers <- FindAllMarkers(CD4_cluster, min.pct = 0.2, assay = "RNA")

top10 <- gt.markers %>%
  group_by(cluster) %>%
  slice_max(n = 10, order_by = avg_log2FC)

top10_heatmap <- DoHeatmap(CD4_cluster, features = unique(top10$gene)) 
top10_heatmap

#ggsave("CTF446_updatedfiles/output/paper/top10_heatmap.png", plot = last_plot(), units = "cm", width = 10, height = 30)
```

##dotplots genes

```{r}

library(SCpubr)

genes1 <- list("Th17" = c(
  "Il17a", "Il17f", "Il22", "Csf2",
  "Rorc", "Stat3","Ccr6", "Il23r"),
              "Th1" = c(
  "Ifng", "Il2", "Tnf","Tbx21", "Stat1", "Stat4", "Runx3"),
              "Th2" = c(
  "Il4", "Il5", "Il13","Gata3", "Stat6"),
              "TFH" = c( "Il21", "Il4",
  "Bcl6", "Maf", "Tox2","Cxcr5", "Pdcd1"),
              "Treg" = c("Foxp3", "Il10", "Tgfb1"),
              "Cytotoxic" = c("Gzmb", "Gzma", "Prf1", "Nkg7", "Fasl", "Klrc1", "Klrc2", "Klrk1", "Cd107a"),
                              "Other" = c("Cd40lg", "Bhlhe40"))




p <- SCpubr::do_DotPlot(sample = CD4_cluster, 
                        features = genes1,
                        sequential.palette = "YlOrRd" )
p

#ggsave("CTF446_updatedfiles/output/paper/CD4subset_scpubr.png", plot = last_plot(), units = "cm", width = 30, height = 10)

```


```{r}
library(RColorBrewer)
genes2 <- list("Memory" = c(
  "Il7r",    # IL-7 receptor, supports survival
  "Tcf7", # Transcription factor driving memory formation
  "Lef1",
  "Bcl2",    # Anti-apoptotic gen
  "Id2", # Regulates differentiation toward memory
  "Id3",
  "Fas",
  "Cd27", 
  "Sell",    # CD62L synonym
  "Ccr7",    # Lymphoid homing, central memory
  "Eomes"   # Long-lived memory transcription factor
  ),
"Activation" = c("Il15ra",
  "Il2ra",    # CD25, involved in memory precursor activation
  "Cd44",    # Memory marker
  "Icos",
  "Cd28",
  "Lag3",
  "Klrg1",
  "Pdcd1",
  "Havcr2",
  "Ctla4"),
              "Trafficking/residency" = c("S1pr1","Cxcr6", "Cxcr3", "Ccr5", "Cd69", "Itga4", "Itgb1", "Itgae", "Itga1", "Itgal", "Notch1","Zfp683", "Prdm1", "Runx3","Ahr", "Crtam", "Cd101"))




p <- do_DotPlot(
  sample = CD4_cluster,
  features = genes1
)

p

q <- SCpubr::do_DotPlot(sample = CD4_cluster, 
                        features = genes2,
                        sequential.palette = "YlOrRd")
q


#ggsave("CTF446_updatedfiles/output/paper/CD4type_scpubr.png", plot = last_plot(), units = "cm", width = 30, height = 10)

```


###renaming cluster id
```{r}
new.cluster.ids <- c("Int. Th", "ISG+", "Tcf7+", "Teff", "Th17")
names(new.cluster.ids) <- levels(CD4_cluster)
CD4_cluster <- RenameIdents(CD4_cluster, new.cluster.ids)

y <- DimPlot(CD4_cluster, reduction = "umap", label = F, label.size = 6) + NoLegend()

y

y<- LabelClusters(
  plot = y,
  id = "ident",
  repel = TRUE,
  size = 6,
  box = TRUE,
  fill = alpha("white", 0.75),  # 75% transparent white
  color = "black"               # black text
)

y <- y + labs(
  x = "UMAP 2",
  y = "UMAP 1") + theme_umap_arrows()
y

#ggsave("CTF446_updatedfiles/output/paper/new_clusterID.png", plot = y, units = "cm", width = 15, height = 10)
#saveRDS(CD4_cluster, "CTF446_updatedfiles/seu2.rds")
```


#5 Top clonotypes

#Clonotype information our of all cells

```{r}

library(SingleCellExperiment)
library(singleCellTK)

colorblind_vector <- hcl.colors(n=7, palette = "inferno", fixup = TRUE)

clonaloverlap <-clonalOverlap(CD4_cluster,
              cloneCall = "strict",
              method = "morisita")

#ggsave("CTF446_updatedfiles/output/paper/clonaloverlap.png", plot = clonaloverlap , units = "cm", width = 15, height = 10)

```

##clonesize

```{r}
y <- DimPlot(CD4_cluster, group.by = "cloneSize", pt.size = 1.5 ) +
    scale_color_manual(values=rev(colorblind_vector[c(1,3,5,7)]))

y <- y + labs(
  x = "UMAP 2",
  y = "UMAP 1") + theme_umap_arrows()
y

#ggsave("CTF446_updatedfiles/output/paper/clonesize_umap.svg", plot = y, units = "cm", width = 15, height = 10)

```



##Frequency of top10 clonotypes
##Top clonotypes


```{r}
###top50_frequency
# 1. Extract clonalProportion column, remove NA
clonal_freq_values <- CD4_cluster@meta.data[["clonalProportion"]]
clonal_freq_values <- na.omit(clonal_freq_values)

# 2. Get unique clonalProportion values
unique_clonal_freq <- unique(clonal_freq_values)

# 3. Sort unique values in descending order
sorted_clonal_freq <- sort(unique_clonal_freq, decreasing = TRUE)

print(sorted_clonal_freq)

# 4. Slice the top 50 values
top_50_clonal_freq <- head(sorted_clonal_freq, 100)

# 5. Print top 50
print(top_50_clonal_freq)

percentage <- (top_50_clonal_freq*100)

# Load ggplot2 for plotting
library(ggplot2)


# Create a data frame for plotting
percentage_df <- data.frame(
  Percentage = percentage
)

# Create the bar plot
perc <- ggplot(percentage_df, aes(x = factor(1:length(percentage)), y = Percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Total 6,726 clones detected",
    x = "All Clonotypes",
    y = "Percent of total clonotypes(%)"
  ) +
  theme(axis.text.x = element_blank())  # Remove x-axis labels

#ggsave("CTF446_updatedfiles/output/paper/clonotype_percentage.png", plot = perc, units = "cm", width = 15, height = 10)

```

## Top6 clonotypes
```{r}
#extract top6 clonotypes, CTstrict equivalent to clonotype
# counts per strict clonotype
ct_counts <- sort(table(CD4_cluster@meta.data$CTstrict), decreasing = TRUE)

# top 6 clonotype names
top6_strict_names <- names(ct_counts)[1:6]

# subset metadata for those exact clonotypes (guarantees 6 clonotype labels)
top6_metadata_strict <- CD4_cluster@meta.data[CD4_cluster@meta.data$CTstrict %in% top6_strict_names, ]

# sanity checks
length(top6_strict_names)            # should be 6
table(top6_metadata_strict$CTstrict) # counts for each of the top 6

# Extract top 6 names as a character vector
top6_CTstrict_vector <- names(ct_counts)[1:6]

top6_CTstrict_vector

#subset top6 clonotypes for scVelo

CD4_top6 <- subset(
  CD4_cluster,
  subset = CTstrict %in% top6_CTstrict_vector
)

#saveRDS(CD4_top6, "Celltypist/seu_top6clones.rds")

```



```{r}
library("APackOfTheClones")


CD4_cluster <- RunAPOTC(CD4_cluster, run_id = "first", clonecall = "strict", override = T)

clone_plot <- APOTCPlot(CD4_cluster, run_id = "first", clonecall = "strict", legend_buffer = -1)

clone_plot


library(viridis)

clones <- showCloneHighlight(clone_plot,
    clonotype = top6_CTstrict_vector,
    color_each = c(red     = "#CD2626",
  orange  = "#FF7F24",
  yellow  = "#FFD700",
  green   = "#00CD00",
  blue    = "#009ACD",
  purple  = "#8B4789"),
  # pink    = "#FF69B4",  
  # brown    = "#8B4513",  
  # beige    = "#FAFAD2",  
  # black  = "#030303" 
    scale_bg = 0.95, 
    fill_legend = F) + theme_umap_arrows()

clones 

#ggsave("CTF446_updatedfiles/output/paper/top6clonotypes.png", plot = clones, units = "cm", width = 20, height = 15)

```

##2.5 Trajectory analysis Monocle
```{r}
library(monocle3)
library(SeuratWrappers)
library(tidyverse)


cds <- as.cell_data_set(CD4_cluster, group.by = "ident")
colData(cds)

plot_cells(cds, color_cells_by = "ident",
           label_groups_by_cluster = FALSE,
           group_label_size = 5) +
  theme(legend.position = "right")


#learn the graph for trajectory inference
#As we have already the UMAP information from the seurat object so we just need to cluster the cells from the umap information from the seurat
cds <- cluster_cells(cds, resolution = 9e-5)

cds <- learn_graph(cds, use_partition = FALSE)

trajectory_mono <- plot_cells(cds, color_cells_by = "cluster",
           label_leaves=FALSE,
           label_branch_points=FALSE,
           trajectory_graph_color = "#030303",
           trajectory_graph_segment_size = 1,
           label_groups_by_cluster = TRUE) + theme_umap_arrows() + 
  theme_void() + 
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )+ theme(legend.position = "none")

trajectory_mono

#ggsave("CTF446_updatedfiles/output/paper/trajectory_mono.png", plot = trajectory_mono, units = "cm", width = 15, height = 10)


```

#Slingshot trajectory

```{r}
library("SeuratExtend")

CD4_cluster <- RunSlingshot(
  CD4_cluster,
  group.by = "SCT_snn_res.0.16",
  reducedDim = "PCA"
)

sling <- CD4_cluster@misc$slingshot$PCA$SlingPseudotime

CD4_cluster@meta.data[,colnames(sling)] <- as.data.frame(sling)

DimPlot2(CD4_cluster, features=colnames(sling), cols = "C", theme = NoAxes())

trajectory_sling <- DimPlot.Slingshot(CD4_cluster, theme = NoAxes()) + theme_umap_arrows()

trajectory_sling

#ggsave("CTF446_updatedfiles/output/paper/trajectoryLine_slingshot.png", plot = trajectory_sling, units = "cm", width = 15, height = 10)

```

